```{r}
# Load libraries
library(ggplot2)
library(dplyr)
library(alr3)
```

```{r}
pf <- read.csv("../datasets/pseudo_facebook.tsv", sep="\t")
```

# Scatter Plots
```{r}
qplot(x = age, y = friend_count, data= pf)

# another way
qplot(age, friend_count, data=pf)
```

# ggplot syntax
```{r}
ggplot(aes(x=age, y=friend_count), data = pf) + geom_point() +
 xlim(13,90)
summary(pf$age)
```

# Overplotting
# Whenever there is toomany points over a region
```{r}
ggplot(aes(x=age, y=friend_count), data = pf) +
  geom_jitter(alpha=1/20) + # transparency of the plot
  xlim(13,90) 

```
# Let's transform the Y axis by using coord_trans
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) +
  geom_point(alpha=1/20, position=position_jitter(h=0)) +
  xlim(13,90) + 
  coord_trans(y = "sqrt")
```
# Alpha and  Jitter
Explore the relationship between friends initiated vs age 
```{r}
ggplot(aes(x=age, y=friendships_initiated), data=pf) +
  geom_jitter(alpha=1/10, position = position_jitter(h=0)) + 
  coord_trans(y="sqrt")
```

# Conditional Means
```{r}
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n())

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

# Another way to do it :) 
```{r}
pf.fc_by_age <- pf %>% 
  group_by(age) %>% 
  summarise(friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n()) %>% 
  arrange(age)

head(pf.fc_by_age, 20)
```

```{r}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) +
  geom_line(position=position_jitter(h=0))
```

#OVerlay summaries with Raw Data
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  coord_cartesian(xlim=c(13,90), ylim=c(0,1000)) + 
  geom_point(alpha=0.05, 
             position=position_jitter(h=0),
             color='orange') + 
  geom_line(stat="summary", fun.y=mean, color="red") + 
  geom_line(stat="summary", fun.y=quantile, fun.args=list(probs=0.1), linetype=2,color="blue") +
  geom_line(stat="summary", fun.y=quantile, fun.args=list(probs=0.5), color="yellow") + 
  geom_line(stat="summary", fun.y=quantile, fun.args=list(probs=0.9), linetype=2,color="black")
```
# Correlation
```{r}
cov_pf = cov(x=pf$age,y=pf$friend_count, method="pearson")
std_x = sd(pf$age)
std_y = sd(pf$friend_count)

r = cov_pf / (std_x * std_y)
r

# using cor.test
cor.test(x = pf$age,y = pf$friend_count, method="pearson")

# using with 
with(pf, cor.test(x = age,y = friend_count, method="pearson"))
```
# Using monotonic functions
```{r}
# using with 
with(data = subset(pf, age<=70), cor.test(x = age,y = friend_count, method="pearson"))
```

# Correlation methods 
# Spearman 
```{r}
with(data = subset(pf, age<=70), cor.test(x = age,y = friend_count, method="spearman"))
```

# Create ScatterPlots 
```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) + 
  coord_cartesian(xlim=c(0,500), ylim=c(0,1000)) + 
  geom_point(alpha=0.05, 
             position=position_jitter(h=0),
             color='orange')
```

# Strong correlations
```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) + 
  geom_point(alpha=0.05, 
             position=position_jitter(h=0),
           color='orange') + 
  xlim(0, quantile(pf$www_likes_received, p=0.95)) +
  ylim(0, quantile(pf$likes_received, p=0.95)) +
  geom_smooth(method = 'lm', color = 'red')
  
  
```   

# Correlation  of www_likes_received & likes_received
```{r}
with(data = pf, cor.test(x = www_likes_received,y = likes_received, method="pearson"))
```


# More caution with correlation 
```{r}
data("Mitchell")
?Mitchell
```

# Scatterplot temp vs months
```{r}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point(alpha=0.1, 
             position=position_jitter(h=0),
           color='black') 
```

# Correlation of Noisy Scatterplots
```{r}
with(data = Mitchell, cor.test(x = Month,y = Temp, method="pearson"))
```

# Making sense of Data 
```{r}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point(alpha=0.1, 
             position=position_jitter(h=0),
           color='black') +
  scale_x_discrete(breaks=seq(0,203,12))

range(Mitchell$Month)
```

```{r}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point()
```

```{r}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) +
  geom_line(position=position_jitter(h=0))

head(pf.fc_by_age, 10)
pf.fc_by_age[17:19,]
```

# Understanding noise
```{r}
pf$age_with_months <- pf$age + (1 - pf$dob_month / 12)
head(pf$age_with_months)
```

# Age with months Means
```{r}
pf.fc_by_age_months <-  pf %>% 
  group_by(age_with_months) %>% 
  summarise(friend_count_mean=mean(friend_count),
          friend_count_median=median(friend_count),
          n=n()) %>% 
  arrange(age_with_months)

head(pf.fc_by_age_months)

```

# Noise in Conditonal Means
```{r}
p1 <- ggplot(aes(x=age, y=friend_count_mean), data=subset(pf.fc_by_age, age < 71)) +
  geom_line(position=position_jitter(h=0)) + geom_smooth()

p2 <- ggplot(aes(x=age_with_months, y=friend_count_mean), data=subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line(position=position_jitter(h=0)) + geom_smooth()

p3 <- ggplot(aes(x=round(age / 5) * 5, y=friend_count), data=subset(pf, age < 71)) +
  geom_line(position=position_jitter(h=0), stat='summary', fun.y=mean)

library(gridExtra)

grid.arrange(p2,p1, p3, ncol=1)
```

