Exploratory Data Analysis by Laura Uzcategui
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(GGally)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityWhites.csv')
```

# Summary 
  The following project for exploratory data analysis is based on a White Wine 
  quality data provided on the Research paper 
  [Modeling Wine Preferences from 
  Physicochemical Properties using Fuzzy Techniques](http://www.scitepress.org/Papers/2015/55519/55519.pdf)
  on which the purpose was to predict the human test preference based on 
  physico chemical properties of wine. 
  
  The data was obtained from a variety of wines with origin in Portugal, 
  specifically "Vinho Verde".

# Univariate Plots Section

Lets explore how the data looks by using the head function, which will show us the first rows. 

```{r echo=FALSE, Univariate_Plots}
head(wine)
```
**Initial Observation**
 - Clearly we can see how all of our data is numeric. 

Lets explore the dimensions of the dataset, which will give us an idea on how 
big is the dataset. 

```{r echo=FALSE, Univariate_Plot1}
dim(wine)
```

Now, lets dig a bit deeper and see the structure of the dataset by using str method

```{r echo=FALSE}
str(wine)
```

**Initial Observation**
 - Data is nott only numerical but integer too.

Last but not least, let's take a look to the Summary of the dataset
```{r echo=FALSE}
summary(wine)
```

**Initial Observation**: initially our data looks good, but I would 
be interested to explore why free and total sulfur dioxide have such a big 
amounts on relation to other variables

Now lets explore with some charts, starting with Alcohol levels

```{r echo=FALSE}
ggplot(aes(x=alcohol), data=wine, ) +
  geom_histogram(color = 'black', fill='#d6e8d5', binwidth = 0.5)
```

**Initial Observation**
 - It seems there is few relevant data points and binwidths we might 
 want to check a  bit further for that let's summarise alcohol variable 
 
```{r echo=FALSE}
summary(wine$alcohol)
```

Let's set some limits to the data that we are observing 
```{r echo=FALSE}
ggplot(aes(x=alcohol), data=subset(wine, !is.na(alcohol))) +
  geom_histogram(color = 'black', fill='#d6e8d5', bins = 30, binwidth = 0.1) + 
  coord_cartesian(xlim=c(8.00,14.20),ylim=c(0,250)) 
```

**Initial Observation**
- Now the data seems to be more uniform and even show a pattern

Let's keep exploring other variables, such as volatic acidity 
which is the amount of acetic acide and might influence taste
```{r echo=FALSE}
ggplot(aes(x=volatile.acidity), data=wine ) +
  geom_histogram(color = 'black', fill='#ffecc9', binwidth = 0.05)
```

Same as before, checking the summary will give a better idea on limits 
for the x axis
```{r echo=FALSE}
summary(wine$volatile.acidity)
```
```{r echo=FALSE}
ggplot(aes(x=volatile.acidity), data=wine ) +
  geom_histogram(color = 'black', fill='#ffecc9', binwidth = 0.03) +
  coord_cartesian(xlim=c(0.0800,1.1000),ylim=c(0,800)) 
```


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
The structure of the dataset is  composed of4898 observations (rows) and 13
variables (columns)

### What is/are the main feature(s) of interest in your dataset?
The main features of interest in my dataset are: 
- Alcohol 
- Volatic Acidity 
- Residual sugar
- Total sulfur dioxide

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
- Sulfates which might contribute to total sulfur dioxide levels

### Did you create any new variables from existing variables in the dataset?
- No. I have not created any new variables initially.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
- I did apply some adjustments of the bins and binwidht on the data, specifically
for the alcohol and volatic.acidity of the wine, since I wanted to check how the 
data was spread out in the distribution.


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!