Exploratory Data Analysis by Laura Uzcategui
========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(GGally)
library(Hmisc)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine <- read.csv('wineQualityWhites.csv')
```

# Summary 
  The following project for exploratory data analysis is based on a White Wine 
  quality data provided on the Research paper 
  [Modeling Wine Preferences from 
  Physicochemical Properties using Fuzzy Techniques](http://www.scitepress.org/Papers/2015/55519/55519.pdf)
  on which the purpose was to predict the human test preference based on 
  physico chemical properties of wine. 
  
  The data was obtained from a variety of wines with origin in Portugal, 
  specifically "Vinho Verde".

# Univariate Plots Section

Lets explore how the data looks by using the head function, which will show us the first rows. 

```{r echo=FALSE, Univariate_Plots}
head(wine)
```
**Initial Observation**
 - Clearly we can see how all of our data is numeric. 

Lets explore the dimensions of the dataset, which will give us an idea on how 
big is the dataset. 

```{r echo=FALSE, Univariate_Plot1}
dim(wine)
```

Now, lets dig a bit deeper and see the structure of the dataset by using str \
method

```{r echo=FALSE}
str(wine)
```

**Initial Observation**
 - Data is nott only numerical but integer too.

Last but not least, let's take a look to the Summary of the dataset
```{r echo=FALSE}
summary(wine)
```

**Initial Observation**: initially our data looks good, but I would 
be interested to explore why free and total sulfur dioxide have such a big 
amounts on relation to other variables

Now lets explore with some charts, starting with Alcohol levels

```{r echo=FALSE}
ggplot(aes(x=alcohol), data=wine, ) +
  geom_histogram(color = 'black', fill='#d6e8d5', binwidth = 0.5)
```

**Initial Observation**
 - It seems there is few relevant data points and binwidths we might 
 want to check a  bit further for that let's summarise alcohol variable 
 
```{r echo=FALSE}
summary(wine$alcohol)
```

Let's set some limits to the data that we are observing 
```{r echo=FALSE}
ggplot(aes(x=alcohol), data=subset(wine, !is.na(alcohol))) +
  geom_histogram(color = 'black', fill='#d6e8d5', binwidth = 0.1) + 
  coord_cartesian(xlim=c(8.00,14.20),ylim=c(0,250)) 
```

**Initial Observation**
- Now the data seems to be more uniform and even show a pattern

Let's keep exploring other variables, such as volatic acidity 
which is the amount of acetic acide and might influence taste
```{r echo=FALSE}
ggplot(aes(x=volatile.acidity), data=wine ) +
  geom_histogram(color = 'black', fill='#ffecc9', binwidth = 0.05)
```

Same as before, checking the summary will give a better idea on limits 
for the x axis
```{r echo=FALSE}
summary(wine$volatile.acidity)
```
```{r echo=FALSE}
ggplot(aes(x=volatile.acidity), data=wine ) +
  geom_histogram(color = 'black', fill='#ffecc9', binwidth = 0.03) +
  coord_cartesian(xlim=c(0.0800,1.1000),ylim=c(0,800)) 
```

**Initial Observation**
- Now the data seems to be more uniform and even show a pattern

We know from a fact that residual sugar might have influence in the levels of 
alcohol, so let's check this variable too
```{r echo=FALSE}
ggplot(aes(x=residual.sugar), data=wine ) +
  geom_histogram(color = 'black', binwidth = 5) 
```

It seems we need to adjust a bit of bidwith and the limits of both axis
```{r echo=FALSE}
summary(wine$residual.sugar)
```

With the maximum and minimum values,lets limit the axis and plot it
```{r echo=FALSE}
ggplot(aes(x=residual.sugar), data=subset(wine, !is.na(residual.sugar))) +
  geom_histogram(color = 'black', binwidth = 0.05) +
  coord_cartesian(xlim=c(0.6,70),ylim=c(0,350)) + 
  scale_x_log10()
  
```

Now, let's check the total sulfure dioxide
```{r echo=FALSE}
ggplot(aes(x=total.sulfur.dioxide), data=wine ) +
  geom_histogram(color = 'black', fill='#253647', bins = 20) 
```

```{r echo=FALSE}
summary(wine$total.sulfur.dioxide)
```

```{r echo=FALSE}
ggplot(aes(x=total.sulfur.dioxide), data=subset(wine, !is.na(total.sulfur.dioxide))) +
  geom_histogram(color = 'black', fill='#253647', binwidth = 5) +
  coord_cartesian(xlim=c(9,440.0),ylim=c(0,280))
  
```

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
The structure of the dataset is  composed of4898 observations (rows) and 13
variables (columns)

### What is/are the main feature(s) of interest in your dataset?
The main features of interest in my dataset are: 
- Alcohol 
- Volatic Acidity 
- Residual sugar
- Total sulfur dioxide

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
- Sulfates which might contribute to total sulfur dioxide levels

### Did you create any new variables from existing variables in the dataset?
- No. I have not created any new variables initially.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
- I did apply some adjustments of the bins and binwidht on the data, specifically
for the alcohol and volatic.acidity of the wine, since I wanted to check how the 
data was spread out in the distribution and noticed difference that makes to 
adjust limits between axis and make the scale different. 


# Bivariate Plots Section

On this section we will explore relationships between two variables, initially
we will start to explore the table of correlation coeficients, then we will 
continue to explore the the main features selected to then explore also the \ 
relationship between those variables that were not selected as main features.

```{r echo=FALSE }
cr <- rcorr(as.matrix(wine))
cr$r
```

**Initial Observations** 
 
 It seems pretty clear that some of the asumptions about relationships between / 
 variables selected as main ( residual.sugar, alcohol, volatic.acidity and 
 total.sulfur.dioxide) might been well chosen as its coefficients show certain 
 correlation. 
 
 Such as: 
 
 - Alcohol - Residual.Sugar --> -0.4506
 - total.sulfur.dioxide - Residual.Sugar --> 0.401
 - Alcohol - total.sulfur.dioxide --> -0.448
 
 Also, it has been observed that there is a tendency on strong correlation on 
 other variables that were not picked up and interestingly enough came out in 
 the table with correlation. Lets add them to our list for exploration
 
 - Density - residual.sugar --> 0.838
 - Density - alcohol --> -0.780
 - Density - total.sulfur.dioxide --> 0.529
 
  Correlation between main features vs. Dependent variable ( Quality )
 
 - Volatic.Acidity - Quality --> -0.19
 - Alcohol - Quality --> 0.435
 
 As we can see above, density seems an interesting strong correlated variable
 with many of the main features already selected, therefore we will include it 
 into the analysis.
 
 With this information we can proceed further in our investigation on how 
 this physio-chemical components might affect or not the quality of the wine 
 and also the components itself might react.

Let's start by plotting first the correlations between Alcohol and Residual 
Sugar 

```{r echo=FALSE }
ggplot(aes(x=alcohol, y=residual.sugar), data = wine) + 
  geom_point(alpha=1/4, color='blue', fill='blue') + 
  xlim(8,15) 
summary(wine$alcohol)
```

It seems that as the alcohol level increases the residual sugar after \
fermentation rarely goes above 20 grams/liter, and also within the 
scatter plot looks a bit of decrease as the alcohol level goes up.

Let's see if by applying geosmooth function we are right in our asumption

```{r echo=FALSE }
ggplot(aes(x=alcohol, y=residual.sugar), data = wine) + 
 geom_line(position=position_jitter(h=0)) + geom_smooth()
```

Something seems still a bit off, lets try to do a bit more work by applying 
a transformation on the y variable, alcohol is being expressed in % meanwhile 
sugar is being expressed as grames per liter 


```{r echo=FALSE }
ggplot(aes(x=alcohol, y=(residual.sugar/1000)*100), data = wine) + 
  geom_line(position=position_jitter(h=0),stat='summary', fun.y=mean) + 
  geom_smooth()
```

Here we can observe a more clear pattern on how the levels of sugar seems to 
decrease as the level of alcohol increase, starting with around 1.5% level and 
progressively and slowly decreasing until 0.5% and less.

Lets proceed further with the analysis for: total.sulfur.dioxide and 
Residual.Sugar

```{r echo=FALSE }
ggplot(aes(x=residual.sugar ,y=total.sulfur.dioxide), data = wine) + 
  geom_point(alpha=1/2, color='#96b585', fill='#96b585') +
  xlim(0.6,66) + 
  ylim(9,440)
summary(wine$residual.sugar)
summary(wine$total.sulfur.dioxide)
```

Even though there might be a strong correlation, it's still not clear based on 
the plot, so let us check how the scales are represented for both axis by 
reviewing the facts.

- We know already residual sugar is grams/per liter. 
- In the other hand, total.sulfur.dioxide is described as the amount of free 
and bounds forms of SO2, which is represented usually in ppm ( Part per million)

Therefore we might want to transform it to a common scale which will be 
ppm to grames/perliter 

```{r echo=FALSE }
ggplot(aes(x=residual.sugar ,y=(total.sulfur.dioxide/1000)), data = wine) + 
  geom_point(alpha=1/3, color='#96b585', fill='#96b585') +
  xlim(0.6,30)
```

Our data is better spread out now that we have adjusted the scales. 
Nevertheless there seems to be a strong positive correlation between both variables
I would be interested to see if we can get a bit more information by plotting 
with conditional means

```{r echo=FALSE}
wine$tt.sulfur.grm.liter <- wine$total.sulfur.dioxide / 1000

wine.sugar_dioxide <- wine %>% 
  group_by(residual.sugar) %>% 
  summarise(total_mean=mean(tt.sulfur.grm.liter),
          total_median=median(tt.sulfur.grm.liter),
          n=n()) %>% 
  arrange(residual.sugar)

```

```{r echo=FALSE}
ggplot(aes(x=residual.sugar, y=total_mean), data=wine.sugar_dioxide) +
  geom_line(position=position_jitter(h=0)) +
  xlim(0.6,30) + geom_smooth(method='lm')
```

In order to do this plot, a new feature has been added, so next plots will be
easily computed. The new feature is above ppm being transformed to grams/liter 

Now lets proceed to explore our main features with other features that it might 
be interesting from what we found in the correlations

 - Density - residual.sugar --> 0.838
 - Density - alcohol --> -0.780
 - Density - total.sulfur.dioxide --> 0.529

```{r echo=FALSE }
ggplot(aes(x=residual.sugar, y=density), data = wine) + 
  geom_point(alpha=1/4) +
  xlim(0.6, 65.8) 
summary(wine$residual.sugar)
```

```{r echo=FALSE }
ggplot(aes(x=alcohol, y=density), data = wine) + 
  geom_point(alpha=1/4) + 
  xlim(8,15) 
summary(wine$alcohol)
```

```{r echo=FALSE }
ggplot(aes(x=total.sulfur.dioxide, y=density), data = wine) + 
  geom_point(alpha=1/4) +
  xlim(9, 440.0) 
summary(wine$total.sulfur.dioxide)
```


# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

I have noticed certain correlations between the main features and the 
the other features that hasn't been picked, such  as density. It appears 
main features such as residual sugar, alcohol and total sulfur dioxide might
affect the density of the wine. 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?
The strongest relationship that has been found is a positive strong relation 
between Residual Sugar and Density with a coefficient of 0.83 and the 
representation of the plots, it is prudent to say that is the strongest relationship 
found along features.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!